---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Loading Data
```{r, echo=T}
# Load knitr library
library(knitr)
# Unzip zipfile, acitivity.zip
unzip(zipfile = "activity.zip")
# Read activity.csv and save it as activity data frame
activity <- read.csv("activity.csv", header=T, stringsAsFactors = F)
```
  
2. Process/transforming the data
```{r, echo=T}
# Transform the date column class date
activity <- transform(activity, date = as.Date(date))
```


## What is mean total number of steps taken per day?
1. Making a histogram of the total number of steps taken each day
```{r, echo=T}
# Get sum of steps taken per day
meanDay <- tapply(activity$steps, activity$date, sum, na.rm=T)
# Print histogram of meanDay
hist(meanDay)
```
  
2. The mean and median total number steps taken per day
```{r, echo=T}
# Get mean and median of total number steps taken per day
mean <- summary(meanDay)[4]
median <- summary(meanDay)[3]
# Print the mean and median
print (mean)
print (median)
```

## What is the average daily activity pattern?
1. Making a time series plot of the 5 minute interval(x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=T}
# Get the mean of steps taken for each interval
daily <- tapply(activity$steps, activity$interval, mean, na.rm=T)
# Convert the daily array to data frame
daily <- as.data.frame(as.table(daily))
# Name daily columns
colnames(daily) <- c("Interval", "AveSteps")
# Transform the interval class as numeric
daily <- transform(daily, Interval = as.numeric(as.character(Interval)))
# Plot a time series plot of the 5-min interval and average number of steps taken averaged across all days
plot(daily$AveSteps ~ daily$Interval, type="l")
```
  
2. The 5-minute interval that contains the maximum number of steps
```{r, echo=T}
# Get the Interval with maximum average number of steps
maxInterval <- daily[which.max(daily[,2]),1]
# Print out the maxInterval
print(maxInterval)
```

## Imputing missing values
1. The total number of missing values in the dataset
```{r, echo=T}
# Calculate the total number of missing values
numNA <- sum(is.na(activity$steps))
# Print out numNA
print (numNA)
```
  
2. The NA steps will be replaced by the mean for that 5-minute interval.
```{r, echo=T}
# GetAveStep function that grabs aveage step with given interval
GetAveSteps <- function(x){
  daily[daily$Interval==as.numeric(x),2]
}
# Use GetAveStep function to fill all the missing values
imputeNA <- activity
imputeNA[is.na(imputeNA$steps),]$steps <- sapply(imputeNA[is.na(imputeNA$steps),]$interval, GetAveSteps)
# Get sum of steps taken per day using imputeNA
meanDay_wo_NA <- tapply(imputeNA$steps, imputeNA$date, sum, na.rm=T)
```
  
3. Make a histogram of the total number of steps taken each day
```{r, echo=T}
# Print histogram of meanDay
hist(meanDay_wo_NA)
```
  
4. Calculate and report the mean and median total number of steps taken per day. 
```{r, echo=T}
# Get mean and median of total number steps taken per day
mean_wo_NA <- summary(meanDay_wo_NA)[4]
median_wo_NA <- summary(meanDay_wo_NA)[3]
# Print the mean and median
print (mean_wo_NA)
print (median_wo_NA)
```

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
# Import lubridate library
library(lubridate)
# Create a day column in imputeNA dataframe
imputeNA$day <- wday(imputeNA$date,label=T)
# Indicate if the date is weekend or weekday
imputeNA$wdwk <- (imputeNA$day=="Sun" |imputeNA$day=="Sat")
imputeNA$wdwk <- factor(imputeNA$wdwk, labels=c("weekend", "weekday"))
```
  
2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r, echo=T}
# Aggregate imputeNA by the mean of steps
imputeMean <- with(imputeNA, aggregate(steps, list(interval = interval, wdwk = wdwk), mean))
# Import lattice library
library(lattice)
# Plot two panels for number of steps over interval: one for weekend and the other for weekday. 
with(imputeMean, xyplot(x~interval|wdwk, main="Number of Steps over Interval", ylab="Number of Steps", xlab="Interval", layout=c(1,2), type='l'))

```
